Accessing the nucleon tensor structure in inclusive deep inelastic scattering

### 2016-07-25 1607 - JMC for g2/Paper/JMC_dis.tex
Added references, modified pictures

The tensor charge is a fundamental property of the nucleon, at
present poorly constrained. It has been estimated in lattice
QCD (see, e.g.,
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015ow}), but
only limited information is available from direct measurements. The
way to extract the tensor charge from experimental measurements requires first
of all the extraction the so-called transversity parton distribution function,
denoted by $h_1^q(x)$ \cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa}. The integral of the transversity distribution
corresponds to the contribution of flavor $q$ to tensor charge.

@@

The transversity distribution is notoriously difficult to access because it is
a chiral-odd function and needs to be combined with a spin-flip mechanism to
appear in a scattering process~\cite{}. Usually, this spin flip is provided by another
nonperturbative distribution or fragmentation function, implying that
transversity cannot be accessed in inclusive DIS,
but only in more complex processes such as semi-inclusive DIS or Drell-Yan~\cite{Ralston:1979ys,Collins:1992xw}.

The only other way to attain spin-flip terms in QED and QCD is taking into
account mass corrections. In fact, it is well known that transversity gives a
contribution to the structure function $g_2$ in inclusive DIS, and in
particular to the violation of the so-called Wandzura--Wilczek relation for
$g_2$. However, this contribution is proportional to the current quark mass
and can be expected to be very small.

We revisit the standard analysis of inclusive DIS taking into account the fact
that on-shell quarks cannot be present in the final state, but they rather
decay into hadrons (ideally, forming jets of hadrons). This is sufficient to modify the structure of the DIS
cut-diagram, even if none of the hadrons is detected in the final
state. For a proper description of this effect, we include ``jet
correlators'' into the analysis. We pay particular attention to ensuring that our
results are gauge invariant. We observe that the inclusion of jet correlators
introduces a new
contribution to the inclusive $g_2$ structure function. This term has the
interesting features that: a) violates the Wandzura--Wilczek relation, b)
potentially violates the Burkhardt--Cottingham sum rule, c) is
proportional to the transversity distribution function multiplied by a
nonperturbative ``jet mass'' parameter, much larger than the mass of light
quarks. 
We provide estimates of this
contribution based on a recent extraction of transversity and show that it
could be very large.

@@

where $F_2^{(0)}$ is the structure function calculated with the handbag diagram sporting a bare quark propagator instead of the jet correlator, and $\xi=2x_B/(1+\sqrt{1+4x_B^2M^2/Q^2})$ with $M$ the nucleon's mass is the Nachtmann scaling variable. (We also omitted the dependence of the structure function on $Q^2$ for clarity of notation). In this paper we limit our attention to effects of order $O(1/Q)$ and therefore can extend the integration to $\mu^2=\infty$. Therefore, the jet function $J_2$ decouples and, thanks to the sum rule \eqref{eq:jetfnsprops}, integrates to 1. One then recovers the conventional result,

@@

\begin{figure}[tbh]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left) and neutron (right) $g_2$ structure function compared to the JAM15 fit of the  $g_2-g_2^{\text{WW}}$ (solid black) \cite{JAM15}. The quark and jet contributions are shown with a dotted red and a dot-dashed green line respectively, with uncertainty bands coming form the Pavia15 fit of the transversity function \cite{Pavia15}. The pure twist-3 contribution calculated by Braun et al. \cite{Braun-et-al} is shown as a dashed blue line (no uncertainty estiamte was provided in teh original reference).
}
\end{center}
\end{figure}

>>>>

The tensor charge is a fundamental property of the nucleon, at
present poorly constrained. It has been estimated in lattice
QCD (see, e.g.,
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015ow}), but
only limited information is available from direct measurements. The
way to extract the tensor charge from experimental measurements requires first
of all the extraction the so-called transversity parton distribution function,
denoted by $h_1^q(x)$~\cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa}. 
The integral of the transversity distribution
corresponds to the contribution of flavor $q$ to the tensor charge.

@@

The transversity distribution is notoriously difficult to access because it is
a chiral-odd function and needs to be combined with a spin-flip mechanism to
appear in a scattering process~\cite{Jaffe:1996zw,Barone:2001sp}. Usually, this spin flip is provided by another
nonperturbative distribution or fragmentation function, implying that
transversity cannot be accessed in inclusive DIS,
but only in more complex processes such as semi-inclusive DIS or Drell-Yan~\cite{Ralston:1979ys,Jaffe:1991kp,Jaffe:1993xb,Collins:1993kk}.

The only other way to attain spin-flip terms in QED and QCD is taking into
account mass corrections. In fact, it is well known that transversity gives a
contribution to the structure function $g_2$ in inclusive DIS (see, e.g.,
\cite{Accardi:2009au} and references therein), and in
particular to the violation of the so-called Wandzura--Wilczek relation for
$g_2$~\cite{Wandzura:1977qf}. However, this contribution is proportional to the current quark mass
and can be expected to be very small.

We revisit the standard analysis of inclusive DIS taking into account the fact
that on-shell quarks cannot be present in the final state, but they rather
decay into hadrons (ideally, forming jets of hadrons). This is sufficient to modify the structure of the DIS
cut-diagram, even if none of the hadrons is detected in the final
state. For a proper description of this effect, we include ``jet
correlators'' into the analysis. We pay particular attention to ensuring that our
results are gauge invariant. We observe that the inclusion of jet correlators
introduces a new
contribution to the inclusive $g_2$ structure function. This term has the
interesting features that: a) violates the Wandzura--Wilczek relation, b)
potentially violates the Burkhardt--Cottingham sum rule, c) is
proportional to the transversity distribution function multiplied by a
nonperturbative ``jet mass'' parameter, probably much larger than the mass of light
quarks. 
We provide estimates of this
contribution based on a recent extraction of transversity and show that it
could be very large.

@@

where $F_2^{(0)}$ is the structure function calculated with the handbag
diagram sporting a bare quark propagator instead of the jet correlator, and
$\xi=2x_B/(1+\sqrt{1+4x_B^2M^2/Q^2})$ with $M$ the nucleon's mass is the
Nachtmann scaling variable. (We also omitted the dependence of the structure
function on $Q^2$ for clarity of notation). In this paper we limit our
attention to effects of order $O(1/Q)$ and therefore can extend the
integration to $\mu^2=\infty$. Therefore, the jet function $J_2$ decouples
and, thanks to the sum rule \eqref{eq:jetfnsprops}, integrates to 1. One then
recovers the conventional result, 

@@

\begin{figure}[tbh]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left) and neutron (right) $g_2$ structure function compared to the JAM15 fit of the  $g_2-g_2^{\text{WW}}$ (solid black) \cite{JAM15}. The quark and jet contributions are shown with a dotted red and a dot-dashed green line respectively, with uncertainty bands coming form the Pavia15 fit of the transversity function \cite{Pavia15}. The pure twist-3 contribution calculated by Braun et al. \cite{Braun-et-al} is shown as a dashed blue line (no uncertainty estiamte was provided in teh original reference).
}
\end{center}
\end{figure}
